\documentclass[12pt]{article}
\begin{document}


\renewcommand{\baselinestretch}{1.2}\rm

% ------------- A very short (party type) version of ------------

\title{\vspace{4.5cm}\Huge The Love of Don Perlimplin\\ and Belisa in the Garden\\
 \vspace{15pt}\normalsize An Erotic Lace-Paper Valentine\\
 In Four Scenes\\
 Chamber Version}
\author{Federico Garc\'\i a Lorca\thanks{Abridged and censored by Oleg Pikhurko,
 9 October, 1999.}}
\date{(1931)}
\maketitle
\thispagestyle{empty}
\newpage


% --------- cast --------------

\newcommand{\propusk}{\hspace{5pt}}
\newcommand{\pe}[1][]{{\sc Perlimplin. {\it #1}}\propusk\rm}
\newcommand{\be}[1][]{{\sc Belisa. {\it #1}}\propusk\rm}
\newcommand{\ma}[1][]{{\sc Marcolfa. {\it #1}}\propusk\rm}
\newcommand{\mo}[1][]{{\sc Mother. {\it #1}}\propusk\rm}
\newcommand{\fisp}[1][]{{\sc First Sprite. {\it #1}}\propusk\rm}
\newcommand{\sesp}[1][]{{\sc Second Sprite. {\it #1}}\propusk\rm}
\newcommand{\te}{\noindent\it}

\section*{Prologue}

\te Before Belisa's Balcony

\pe Yes?

\ma Yes.

\pe But why ``yes''?

\ma Just because yes.

\pe And if I should say no?

\ma[acidly] No?

\pe No.

\ma Tell me, Master, the reason for the ``no.''

\pe You tell me, you persevering domestic, the reasons for that
``yes.''

\te Pause.

\ma Twenty and twenty are forty...

\pe[listening] Proceed.

\ma And ten, fifty.

\pe Go ahead.

\ma At fifty years one is no longer a child.

\pe Of course!

\ma I may die any minute.

\pe Good Lord!

\ma[weeping] And what will happen to you all alone in the world?

\pe What will happen?

\te Pause.

\ma That's why you have to marry.

\pe[distracted] Yes?

\ma[sternly] Yes.

\pe[miserably] But Marcolfa... why ``yes''? When I was a child a woman
strangled her husband. He was a shoemaker. I can't forget it. I've
always said I wouldn't marry. My books are enough for me. What good
will marriage do me?

\ma Marriage holds great charms...

\be[within, singing] Ah love, ah love...

\ma My master will see the reason I have.

\pe[scratching his head] She sings prettily.

\ma She is the woman for my master. The fair Belisa.

\pe Belisa... but wouldn't it be better... ?

\ma No. Now come.

\te She takes him by the hand and goes toward the balcony.

\ma Say, ``Belisa.''

\pe Belisa...

\ma Louder.

\pe Belisa!

\te Belisa appears, resplendent in her loveliness.

\be Who calls?

\ma[hides behind the curtains] Answer!

\pe[trembling] I was calling.

\be Yes?

\pe Yes.

\be But why, ``yes''?

\pe Just because yes.

\be And if I should say no?

\pe I would be sorry, because.. we have decided that I want to marry.

\be[laughs] Marry whom?

\pe You.

\be[serious] But... Mama! Mama-a-a!

\ma This is going well.

\te Enter the Mother wearing a great wig.

\be Don Perlimplin wants to marry me. What must I do?

\mo The very best of afternoons to you, my charming little neighbour. I
always said to my poor little girl that you have the grace and
elegance of that great lady who was your mother, whom I did not have
the pleasure of knowing.

\pe Thank you.

\ma[furiously, from behind the curtain] I have decided that we are
going...

\pe We have decided that we are going...

\mo To contract matrimony. Is that no so?

\pe That is so.

\be But, Mama, what about me?

\mo You are agreeable, naturally. Don Perlimplin is a fascinating
husband.

\pe I hope to be one, madam.

\ma[to Don Perlimplin] This is almost settled.

\pe Do you think so?

\te They whisper together.

\mo[to Belisa] Don Perlimplin has many lands. On these are many geese
and sheep. The sheep are taken to market. At the market they give
money for them. Money produces beauty... and beauty is sought after by
all men.

\pe Then...

\mo Ever so thrilled... Belisa... go inside. It isn't well for a
maiden to hear certain conversations.

\be Until later.

\te Belisa leaves.

\mo She is a lily. You've seen her face? But if you should see
further! Just like sugar. But, pardon. I need no call these things to
the attention of a person as modern and competent as you...

\pe I don't know how to express our gratitude.

\mo Oh, ``our gratitude.'' What extraordinary delicacy! The gratitude
of your heart and yourself... I have sensed it...

\ma The wedding.

\pe The wedding...

\mo Whenever you wish. Though... to every mother... until later!

\te Leaves.

\ma At last!

\pe Oh, Marcolfa, Marcolfa! Into what world are you going to thrust
me?

\ma Into the world of matrimony...

\be[inside] Ah love, ah love...

\ma Beautiful maiden.

\pe Like sugar... white inside. Will she be capable of strangling me?

\ma Woman is weak if frightened in time.

\be[inside] Ah love, ah love...

\section*{Scene 1}

\te Don Perimplin's room. Don Perlimplin and Marcolfa. Belisa looks
out of a window.

\ma May my master have a good wedding night!

\pe Good night, Marcolfa.

\te Marcolfa leaves.

\be[to outside] Whoever seeks me ardently will find me. My thirst is
never quenched...

\pe Do I disturb you?.. Are you sleepy?

\be[ironically] Sleepy?

\pe The night has become a little chilly.

\te Rubs his hands. Pause.

\pe I wanted to tell you something.

\be And that is?

\pe I have been late in deciding... but...

\be Say it.

\pe Belisa, I love you.

\be Oh, you little gentleman! That's your duty.

\pe Yes?

\be Yes.

\pe But why ``yes''?

\be[coyly] Because.

\pe Do I have your permission to remove my coat?

\be[Yawning] Of course, little husband. And put out the lights, if
that is your wish...

\te Two sprites run a curtain of misty gray.

\sesp The soul of Perlimplin, tiny and frightened like a newborn
duckling, becomes enriched and sublime at these moments.

\te The sprites laugh.

\fisp The audience is impatient.

\sesp Don Perlimplin, do we help or hinder you?

\fisp Help: because it is not fair to place before the eyes of the
audience the misfortune of a good man.

\sesp Tomorrow the whole world will know about it.

\te They open the curtain. Don Perlimplin appears with two enormous
horns.

\pe[awakening] Belisa! Belisa! Answer me!

\be[pretending to awaken] Perlimplinpinito... what do you want?

\pe Tell me quickly. Why are the balconies open?

\be Because this night the wind has blown as never before.

\pe Why do the balconies have five ladders that reach to the ground?

\be Because that is the custom in my mother's country.

\pe Belisa! Belisa! And why not? You explain everything so well. I am
satisfied. Why couldn't it have been like that?

\be[coyly] I'm not a little fibber.

\pe For the first time in my life I am happy! Ah love, ah love...

\section*{Scene 2}

\te Don Perlimplin's dining room. Don Perlimplin and Marcolfa.

\pe Marcolfa, why do you keep on crying?

\ma Yours Grace knows. On your wedding night five men entered your
bedroom through the balconies. Five! Representatives of the five races
of the earth. The European, with his beard---the Indian---the
Negro---the Yellow Man---and the American. And you unaware of it all.

\pe That is of no importance.

\ma Just imagine: yesterday I saw her with another one.

\pe[intrigued] Really?

\ma My master loves her too much.

\pe Not as much as she deserves.

\ma Here she comes.

\pe Please leave.

\te Marcolfa leaves. Belisa enters.

\be[not noticing Perlimplin] Again I have failed to see him. In my
walk through the park they were all behind me except him...

\pe Ahem!

\be[turning] Oh! What a fright you gave me.

\pe[approaching her affectionately] I observe you were speaking to
yourself.

\be[distastefully] Go away!

\pe Shall we take a walk?

\be No.

\pe Shall we go to the confectioner's?

\be I said no!

\pe Pardon. Poor Belisa! I can see things. And even though it wounds
me deeply, I understand you live in a drama... I know that you are
faithful to me, and that you will continue to be so.

\be[fondly] I've never known any man other than my Perlimplinillo.

\pe That's why I want to help you as any good husband should when his
wife is a model of virtue... Look.

\te He closes the door and adopts a mysterious air.

\pe I know everything! I realized immediately... Has he come by here
today?

\be Twice.

\pe And has he signaled to you?

\be Yes. I haven't seen his face. But...

\pe Don't be afraid to speak to me. I know you love him...

\be He writes me letters.

\pe I know that.

\be But he doesn't let me see him.

\pe That's strange.

\be But there's no doubt he loves me as I wish...

\pe Who could that beautiful youth be?

\be No one know.

\pe[inquisitive] No one?

\be I have asked all my friends.

\pe And if I should tell you I know him?

\be Is that possible?

\pe Wait. Since I am an old man, I want to sacrifices myself to
you. This that I do no one ever did before... Good-bye.

\be Where are you going?

\pe[at the door, grandiosely] Later you will know everything. Wait
till the night. 


\section*{Scene 3}

\te A grove of cypresses and orange trees. Don Perlimplin and
Marcolfa.

\ma Even though I'm not very good at these things, I told her that the
master had instructed me to say... that that young man... would come
tonight at twelve o'clock sharp to the garden, wrapped, as usual, in his
white cape.

\pe And she?

\ma She became as red as a geranium, put her hands to her heart, and
kissed her lovely braids passionately.

\pe[enthusiastic] So she got red as a geranium, eh? And, what did she
say?

\ma She just sighed; that's all. But, oh! such a sigh!

\pe Oh, yes! As no woman ever sighed before! Isn't that so?

\ma Her love must border on madness.

\pe[vibrantly] That's is! What I need is for her to love that youth
more than her own body. And there is no doubt that she loves him.

\ma[weeping] It frightens me to hear you... but how is it possible?
Don Perlimplin, how is it possible that you yourself should encourage
your wife in the worst of sins?

\pe Because Perlimplin has no honour and wants to amuse himself! Now
do you see?

\ma Let my master know that from this moment I consider myself
dismissed from his service. We servants also have a sense of shame.

\pe Oh, innocent Marcolfa! Wait until tomorrow. Now go and perform
your duty. You will do what I have told you?

\ma[leaving, drying her tears] What else is there for me to do? What
else?

\te Don Perlimplin waits till Marcolfa leaves, cautiously looks around,
gets a white cape from a secret recess and puts it on.

\pe Ah love, ah love...

\te Leaves.



\end{document}
